---
title: "Frothy Head Brewing Analysis"
author: "Daniel M. Turner, Pankaj Kumar, Kay Ayala"
date: "December 17, 2018"
output:
  html_document:  
    toc: true    
    theme: united 
params: 
  beer_file_path: ./Analysis/data/Beers.csv
  brewery_file_path: ./Analysis/data/Breweries.csv
  echo: FALSE
  kable_align: 'l'
  
#Please install!!
#install.packages("devtools")
#devtools::install_github("slowkow/ggrepel")

  
---


In response to Frothy Head Brewing’s inquiry on beer and breweries, we have examined the provided datasets and produced metrics and visualizations in order to gain insight into beer production in the United States. 

We have provided a table and a map of how many breweries there are in each state. From these we can get an idea of the distribution of breweries across the country.


```{r echo =params$echo }
library(knitr)
beer <- read.csv(params$beer_file_path, header = TRUE, sep = ",")
brewery <- read.csv(params$brewery_file_path, header = TRUE, sep = ",")

```

```{r echo =params$echo }

# for data cleaning later

#     Change Brewery_Name and Beer_Name to describe the right columns
#     Change Brewery_id and Beer_ID to be the same capitalization
#     Scrub duplicate entries
#     Check for any state abrevations not in the state abreviation list and verify all are capitalized correctly 
#     

```


```{r include =params$echo}
#install.packages("devtools")
#devtools::install_github("slowkow/ggrepel")

##count by state
library(kableExtra)
library(dplyr)
library(usmap)
library(ggplot2)
library(ggrepel)
library(stringr)

my_plot_us= function (regions = c("states", "state", "counties", "county"), 
                      include = c(), data = data.frame(), values = "values", theme = theme_map(), 
                      lines = "black", labels = FALSE, label_color = "black") 
{
  if (!requireNamespace("ggplot2", quietly = TRUE)) {
    stop("Please install `ggplot2`. Use: install.packages(\"ggplot2\")")
  }
  if (!requireNamespace("ggrepel", quietly = TRUE)) {
    stop("Please install `ggrepel`. Use: install.packages(\"devtools\") \n devtools::install_github(\"slowkow/ggrepel\")")
  }
  regions_ <- match.arg(regions)
  if (nrow(data) == 0) {
    map_df <- us_map(regions = regions_, include = include)
    polygon_layer <- ggplot2::geom_polygon(ggplot2::aes(x = map_df$long, 
                                                        y = map_df$lat, group = map_df$group), colour = lines, 
                                           fill = "white", size = 0.4)
  }
  else {
    map_df <- map_with_data(data, values = values, include = include)
    polygon_layer <- ggplot2::geom_polygon(ggplot2::aes(x = map_df$long, 
                                                        y = map_df$lat, group = map_df$group, fill = map_df[, 
                                                                                                            values]), colour = lines, size = 0.4)
  }
  if (labels) {
    if (regions_ == "county" | regions_ == "counties") {
      warning("`labels` is currently only supported for state maps. It has no effect on county maps.")
      label_layer <- ggplot2::geom_blank()
    }
    else {
      centroid_labels <- utils::read.csv(system.file("extdata", 
                                                     paste0("us_", regions_, "_centroids.csv"), package = "usmap"), 
                                         colClasses = c("numeric", "numeric", "character", 
                                                        "character", "character"), stringsAsFactors = FALSE)
      centroid_labels=centroid_labels[order(centroid_labels$abbr),]
      centroid_labels=merge(centroid_labels, data, by.x='fips', by.y='fips')
      if (length(include) > 0) {
        centroid_labels <- centroid_labels[centroid_labels$full %in% 
                                             include | centroid_labels$abbr %in% include | 
                                             centroid_labels$fips %in% include, ]
      }
      label_layer <- ggrepel::geom_text_repel(data = centroid_labels, hjust = 0.5, point.padding = NA,
                                               position = position_dodge(width=0.3),  size=3,
                                        ggplot2::aes(x = centroid_labels$x, y = centroid_labels$y, 
                                                     label = paste(centroid_labels$abbr, centroid_labels$brewsByState)), colour = label_color)
    }
  }
  else {
    label_layer <- ggplot2::geom_blank()
  }
  ggplot2::ggplot(data = map_df) + polygon_layer + label_layer + 
    ggplot2::coord_equal() + theme
}


theme_map <- function(base_size = 9, base_family = "") {
  elementBlank = ggplot2::element_blank()
  `%+replace%` <- ggplot2::`%+replace%`
  unit <- ggplot2::unit
  
  ggplot2::theme_bw(base_size = base_size, base_family = base_family) %+replace%
    ggplot2::theme(axis.line = elementBlank,
                   axis.text = elementBlank,
                   axis.ticks = elementBlank,
                   axis.title = elementBlank,
                   panel.background = elementBlank,
                   panel.border = elementBlank,
                   panel.grid = elementBlank,
                   panel.spacing = unit(0, "lines"),
                   plot.background = elementBlank,
                   legend.justification = c(0, 0),
                   legend.position = c(0, 0))
}


brewsByState = summary(brewery$State)
state_brew= data.frame(brewsByState, check.names = FALSE, stringsAsFactors = TRUE)
state_brew$state=row.names(state_brew)
state_brew$fips = sapply(as.character(str_trim(state_brew$state)), fips)

```

###  Breweries per State


```{r echo =params$echo}

kable(list( state_brew[c(1:17), c('state','brewsByState')], state_brew[c(18:34), c('state','brewsByState')], state_brew[c(35:50), c('state','brewsByState')]), align = params$kable_align, padding = 6L, row.names = FALSE , col.names =c('State__________', 'Breweries')) %>% kable_styling()


```

#### Statewise Brewery Map 
```{r echo =params$echo}

my_plot_us(data = state_brew, values = "brewsByState", label_color = "black", lines = "#1b2f47", labels = TRUE)+
  scale_fill_continuous(name= "number of breweries" , low = "#6bb0f1", high = "#1b2f47")+
    theme(legend.position = "right")
     

```

### Dataset

In order to get an idea of the data in the provided datasets we show the head and tail of the merged data. 

##### Dataset Head

```{r echo =params$echo}
##data merge and check.clean data?

colnames(brewery)[colnames(brewery)=="Brew_ID"] <- "Brewery_id"
brewNBeer <- merge(beer, brewery, by = "Brewery_id" , suffixes = c('_beer', '_brewery'))
names(brewNBeer)= c("Brewery_ID", "Beer_Name", "Beer_ID", "ABV", "IBU",  "Style", "Ounces", "Brewery_Name", "City", "State" )
kable(head(brewNBeer),align = params$kable_align) %>% kable_styling( position = "left")



```

##### Dataset Tail

```{r echo =params$echo}

kable(tail(brewNBeer), align = params$kable_align) %>% kable_styling(bootstrap_options = "striped", full_width = F) 


```

### Missing Values

It is important to note there are missing values in the data and IBU has the most missing values. This is the sum of missing values for each variable. 

```{r echo =params$echo}

##after clean, only missing bitterness? beer types?
kable(colSums(is.na(brewNBeer)),align = params$kable_align, col.names = c('Number of Missing Values')) %>% kable_styling()

# Loading necessary packages for graphing
library(RColorBrewer)

columnNames = colnames(brewNBeer)
columnNames[columnNames=="Brewery_ID"] <- "Brewery ID"
columnNames[columnNames=="Beer_Name"] <- "Beer Name"
columnNames[columnNames=="Beer_ID"] <- "Beer ID"
columnNames[columnNames=="Brewery_Name"] <- "Brewery Name"

missingValues = colSums(is.na(brewNBeer))


# Plots a bar chart for missing values in each variable
print(ggplot() +
  aes(x= columnNames, y=missingValues) + 
  geom_col(width = 0.9, show.legend = TRUE, 
           color= "#89b0d0", fill = "#89b0d0") +
  ggtitle("Number of Missing Values per Variable") +
  theme(plot.title = element_text(hjust = 0.5)) + 
  ylab("Variable") +
  xlab("Number of Missing Values") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
  )

```

### Alcohol Content Distribution by State

In order to get an idea of what alcohol content and bitterness preferences are by state we provide bar charts of ABV and IBU. This may be of use when considering what type of beers to be making in each state. 

```{r echo =params$echo}
library(ggplot2)

##find medians by state. Do twice? once counting NA and once not counting NA
alcy<-tapply(brewNBeer$ABV, INDEX=brewNBeer$State, FUN=median)

#remove na for data
alcy<-tapply(brewNBeer$ABV[!is.na(brewNBeer$ABV)], INDEX=brewNBeer$State[!is.na(brewNBeer$ABV)], FUN=median)

df_abv <- data.frame(state=names(alcy),abv = alcy)

ggplot(data=df_abv, aes(x=state, y=abv, fill = abv ))+
    geom_col()+
    coord_cartesian(ylim=c(0.0375,0.0625))+
    scale_fill_gradient(low="#1b2f47",high="#6bb0f1")+ 
    theme(axis.text.x=element_text(angle=80, hjust=1))+
    guides(fill=FALSE)+
    labs(x = "State", y = "Alcohol Content ABV")

```


### Bitterness Distribution by State

```{r echo =params$echo}
##find medians by state. Do twice? once counting NA and once not counting NA
bitter<-tapply(brewNBeer$IBU, INDEX=brewNBeer$State, FUN=median)

#remove na for data
bitter<-tapply(brewNBeer$IBU[!is.na(brewNBeer$IBU)], INDEX=brewNBeer$State[!is.na(brewNBeer$IBU)], FUN=median)

bitter[is.na(bitter)] <- 0
df_ibu <- data.frame(state=names(bitter),ibu = bitter)
ggplot(data=df_ibu, aes(x=state, y=ibu, fill = ibu)) +
    geom_col()+
    coord_cartesian(ylim=c(15,63))+
    scale_fill_gradient(low="#1b2f47",high="#6bb0f1")+ 
    theme(axis.text.x=element_text(angle=80, hjust=1))+ 
    guides(fill=FALSE)+
    labs(x = "Bitterness IBU", y = "State")


```

### State with the Most Alcoholic ABV Beer and Spread of Alcohol Content ABV

The state with the most alcoholic beer was Colorado. 

```{r echo =params$echo}
##find biggest numbers. NBD
state_data=brewNBeer[which.max(brewNBeer$ABV),]
state_data$State= cell_spec(state_data$State, "html",background  = "red")
kable(format = "html", escape = F, state_data,align = params$kable_align) %>% kable_styling() #alcohol
```


The most produced beers follow the following spread of alcohol content summary. It would be most recommended to produce beer within the first and third quartiles of alcohol content.

```{r echo =params$echo}
##summarize numbers
abv_summary=summary(brewNBeer$ABV)
abv_summary
boxplot(brewNBeer$ABV,
        main = "Alcohol Content ABV in US Beers",
        xlab = "Alcohol Content ABV",
        col = "#758E6E",
        border = "#758E6E",
        horizontal =TRUE)

```

### State with the Most Bitter IBU Beer and Spread of Bitterness Levels IBU

The state with the most bitter beer was Oregon.

```{r echo =params$echo}
state_data=brewNBeer[which.max(brewNBeer$IBU),]
state_data$State= cell_spec(state_data$State, "html",background  = "red")
kable(format = "html", escape = F,state_data ,align = params$kable_align) %>% kable_styling() #bitterness

abv_summary=summary(brewNBeer$IBU)
abv_summary

boxplot(brewNBeer$IBU,
        main = "Bitterness IBU in US Beers",
        xlab = "Bitterness IBU",
        col = "#758E6E",
        border = "#758E6E",
        horizontal =TRUE)

```


### Relationship between the Bitterness and Alcoholic Content of Each Beer

We looked to see if we could find a relationship between bitterness and alcohol content by producing a scatterplot. More alcoholic beers had a the full range of bitterness, but less alcoholic beers tended to have low bitterness. The data has a moderate linear relationship as can be seen from the Pearson's Correlation value and the line on the scatter plot. 

##### Pearson Correlation between ABV and IBU

``` {r echo =params$echo}
cor(brewNBeer$ABV,brewNBeer$IBU, use ="na.or.complete", method="pearson")
  
``` 


```{r echo =params$echo, warning=FALSE}
##plot ibu vs abv. nbd

#setwd('/Users/pankaj/dev/git/smu/Brewery-Analysis')
#state_regions =read.csv('./us_state_region.csv')
#brewNBeer$State=as.character(brewNBeer$State)
#state_regions$State.Code=as.character(state_regions$State.Code)
brewNBeer$State=str_trim(as.character(brewNBeer$State) )
#state_regions$State.Code=str_trim(as.character(state_regions$State.Code) )
#region_bb = merge(brewNBeer, state_regions, by.x= 'State' , by.y = 'State.Code')

ggplot(brewNBeer, aes(x=ABV , y=IBU))+
  geom_point(mapping= aes(color = State))+
  theme_classic()+
  stat_smooth(method = lm)
        #facet_wrap(~ Division)

```
